Analysis and Figures for Gallego et al, 2020

### 2020-09-09 Scripts/Manuscript.Main.Analysis.Rmd
Updated at Wed

plot(layout5)
p1 + 
  p2 +
  # guide_area()+ 
  p3 + plot_layout(design = layout5) + plot_annotation(tag_levels = "A")  -> plot

@@

@@

#overall change in richness
richness.by.point %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
#diatom change in richness
richness.by.phylum.by.point %>%
  filter(phylum == "Bacillariophyta") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
richness.by.phylum.by.point %>%
  filter(phylum == "Dinophyceae") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))

@@

richness.by.phylum.by.point %>% 
  ungroup() %>% 
  dplyr::select(Area, Temperature, pH, Year) %>% 
  distinct() %>% 
  filter(Area == "Hood Canal") %>% 
  mutate(Temperature = 14.995 + 4.299*Temperature,
         pH = 8.2157 + 0.1813*pH) %>% 
  group_by(Year) %>% 
  add_count() %>% 
  summarise(condition = sum(Temperature > 12 & pH < 7.9),
            totalcount = n()) %>% 
  mutate(percent_time = condition / totalcount)

#we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal") %>% 
                    pull(richness))

@@

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island") %>% 
                    pull(richness))

@@

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness))

#while the changes on the San Juan Island are undetectable (p $=$ 0.8).

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness))

@@

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
                    pull(richness))

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness))

@@

Because we are comparing two 
```{r}

## GLOBAL TEST

richness.by.point %>% 
  nest(-Area) %>% 
  mutate(wilcoxon = map(data, ~ wilcox.test(richness ~ Year, data = .x)),
         tidy_wilcox = map_df(wilcoxon, broom::tidy))

## We cannot

richness.by.phylum.by.point %>% 
   nest(-phylum,-Area) %>% 
   mutate(wilcoxon = map (data, ~ wilcox.test(richness ~ Year, data = .x) ),
          wiloc.tidy= map(wilcoxon, broom::tidy))  -> results.lm

results.lm %>%  filter (phylum == "Bacillariophyta") %>% unnest(wiloc.tidy)

results.lm %>%  filter (phylum == "Dinophyceae") %>% unnest(wiloc.tidy)

results.lm %>% 
  unnest(wiloc.tidy) %>%

  ggplot(aes(x = phylum , y = `p.value`)) +
  geom_point(aes(color = Area))
```

>>>>

p1 + 
  p2 +
  # guide_area()+ 
  p3 + plot_layout(design = layout5) + plot_annotation(tag_levels = "A")  -> plot

@@

```{r show change, echo = F}
explore.phylum <- function(phyl){

richness.by.phylum.by.point %>%
  filter(phylum == phyl) %>% 
  group_by(Area, Year) %>% 
  summarise (grp.median = median (richness)) -> medians

richness.by.phylum.by.point %>%
  filter(phylum == phyl) %>% 
  ggplot(aes())+
  geom_histogram(bins = 100, aes(x =  richness, group = Year, color = Year, after_stat(ndensity)),fill = "white", position = "identity") +
  geom_vline(data = medians, aes(xintercept = grp.median, color = Year), linetype = "dashed") +
  facet_wrap (~Area)

}
```

@@

```{r}
#overall change in richness
richness.by.point %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
#diatom change in richness
richness.by.phylum.by.point %>%
  filter(phylum == "Bacillariophyta") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
explore.phylum("Bacillariophyta")

richness.by.phylum.by.point %>%
  filter(phylum == "Dinophyceae") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
explore.phylum("Dinophyceae")

@@

richness.by.phylum.by.point %>% 
  ungroup() %>% 
  dplyr::select(Area, Temperature, pH, Year) %>% 
  distinct() %>% 
  filter(Area == "Hood Canal") %>% 
  mutate(Temperature = 14.995 + 4.299*Temperature,
         pH = 8.2157 + 0.1813*pH) %>% 
  group_by(Year) %>% 
  add_count() %>% 
  summarise(condition = sum(Temperature > 12 & pH < 7.9),
            totalcount = n()) %>% 
  mutate(percent_time = 100 *condition / totalcount)
```

In general, the prevalence of cluster 3 in the Hood Canal will increase from 54.2% of the time to 75% of the time,
```{r echo = F}
# Add the cluster information to the projected scenario

expanded.future %>%

  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
  select(Area, Year, Temperature = TemperatureCelsius,
         pH_new = pH_units) %>% 
  distinct() %>% 
    group_by(Area) %>% 
    nest() %>% 
    left_join(clusters) -> expand.with.clusters

# Project which one on each environmental simulation
expand.with.clusters %>% 
    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
              ) ) %>% unnest(whichComm.res) -> expand.with.clusters
# Calculate the prevalence of each cluster
expand.with.clusters %>%
  group_by(Area, Year, comm) %>% 
  tally() %>%  
  group_by(Area, Year) %>% 
  add_count() %>%
  mutate(prev = n/nn) %>%
  select(Area, Year, comm, prev) %>% pivot_wider(names_from = comm, values_from = prev)
#Note that clusters are different on each Area - cluster #3 from Hood Canal is different from cluster #3 from the San Juans
```

```{r echo = F}
#we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal") %>% 
                    pull(richness), conf.int = T)

@@

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island") %>% 
                    pull(richness), conf.int = T)

@@

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness), conf.int = T)

#while the changes on the San Juan Island are undetectable (difference in median switch of 0.05, with  p $=$ 0.01).

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness), conf.int = T)

@@

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
                    pull(richness),
            conf.int= TRUE)

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
                    pull(richness),
            conf.int= TRUE)

@@

# Responses in diatoms are not all in the same direction - Pseudonitzschia spp shows higher probability of presence at low pH values, and low pH values are more common in the 2095 scenario

expanded.future %>% 
  group_by(Area) %>% 
  filter(  str_detect(taxa, "Pseudo-nitzschia") ) %>% 
  ggplot(aes(x = pH_units, y = present)) +
  geom_smooth(aes(color = as.factor(Year), group = Year)) +
  facet_grid(~Area) + 
  ggtitle("Pseudo-nitzschia spp. response to pH")

expanded.future %>%

  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
  select(Area, Year, Temperature = TemperatureCelsius,
         pH_new = pH_units) %>% 
  distinct() %>% 
    group_by(Area) %>% 
    nest() %>% 
    left_join(clusters) %>% 
    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
              ) ) %>% unnest(whichComm.res)
```
